---
title: "Automatic Conclusion"
author: "Guess Who"
date: "2024-04-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Contexte

```{r lib, echo=FALSE}
library(ggplot2)
library(data.table)
library(gt)
library(gtsummary)
library(GGally)
library(corrplot)
library(Hmisc)
```

Good idea to let some notes about the aim of the report and the methods used for each part.

**Asso tests** part shows how to generate a "gt" summary table and then use the table to automatically adapt the conclusion.

**Correlation tests 2 by 2** part shows how to generate a table of correlation tests for all variables 2 by 2, and then use the table to automatically adapt the conclusion.

## Asso tests

```{r test1, echo=FALSE}
my_data <- iris
my_data$fake.variable <- sample(1:10, nrow(my_data), replace = T) / 5

dt_quanti_vars <- data.table(
  variable = names(my_data),
  is_numeric = unlist(lapply(X = names(my_data), function(x) is.numeric(my_data[[x]]))),
  variable_name = unlist(lapply(X = names(my_data), function(x) gsub("\\.", " ", x)))
)
# dt_quanti_vars

my_group <- "Species"

gt_res <- my_data %>%
  tbl_summary(
    include = all_of(c(dt_quanti_vars$variable[dt_quanti_vars$is_numeric])),
    by = all_of(my_group),
    statistic = list(
      all_continuous() ~ "{mean}", # if quanti
      all_categorical() ~ "{p}%" # if quali
    ),
    missing_text = "Missing"
  ) %>%
  bold_labels() %>%
  add_overall() %>%
  add_p(
    pvalue_fun = scales::label_pvalue(accuracy = .0001)
  ) %>%
  modify_caption(paste0("Table: ",my_group, " vs quantitatif variables selected"))

# show it
gt_res

cat("\n")

## automatic comment : 
tab_res <- gt_res$table_body
for (i in seq_len(nrow(tab_res))) {
  print(
    paste0(
      "- ", tab_res[i, "variable"], 
      ifelse(tab_res[i, "p.value"] < 0.05, " do", " do not"),
      " correlates with ", my_group, " (p = ", formatC(tab_res[[i, "p.value"]]), ")",
      "\n"
    )
  )
}

cat("\n")

## Or only make comment for significant asso : 
tab_res <- gt_res$table_body[gt_res$table_body$p.value<0.05, ]
cat(
  paste0(
    "Following variable are associated with ", my_group, " (p<0.05) : \n",
    paste0(
      dt_quanti_vars$variable_name[dt_quanti_vars$variable %in% tab_res$variable],
      collapse = ", "
    )
  )
)


```


## Correlation tests 2 by 2 


```{r test2, echo=FALSE}
my_data <- iris
my_data$fake.variable <- sample(1:10, nrow(my_data), replace = T) / 5

dt_quanti_vars <- data.table(
  variable = names(my_data),
  is_numeric = unlist(lapply(X = names(my_data), function(x) is.numeric(my_data[[x]]))),
  variable_name = unlist(lapply(X = names(my_data), function(x) gsub("\\.", " ", x)))
)
# dt_quanti_vars
my_group <- "Species"

## graphic ok
ggpairs(my_data, aes(color = my_group))
# think to add a legend about "***" meaning

cat("\n")

## other graphic ok 
corrplot(
  cor(my_data[, c(dt_quanti_vars$variable[dt_quanti_vars$is_numeric])]),
  method = "number",
  type = "upper", # show only upper side
  ## --here to see other params... # p.mat #sig.level 
)

cat("\n")

## table : just cor val
cor_tab <- cor(my_data[, c(dt_quanti_vars$variable[dt_quanti_vars$is_numeric])]) # if wanted
# cor_tab

## Hmisc
res_corr <- rcorr(
  as.matrix(my_data[, c(dt_quanti_vars$variable[dt_quanti_vars$is_numeric])]),
  type = "pearson" # or "spearman" # --here you choose linear test or not
)

# get list of pairs significantly correlated : 
# res_corr
# res_corr$P
# res_corr$P < 0.05

correlated_variables <- data.table(
  variable = row.names(res_corr$P), 
  significantly_correlated_to = unlist(lapply(
    X = row.names(res_corr$P),
    function(ivar) {
      tmp <- res_corr$P[ivar, ]
      paste0(names(which(tmp < 0.05)), collapse = ",")
    }
  ))
)

gt::gt(correlated_variables) |>
  tab_header(
    title = md("Table of **significant** correlation"),
    subtitle = md("`pearson` type of correlation")
  )
# I like a synthetic table. easy to read.  

cat("\n")

# but if you prefer sentences, you can write : 
correlated_variables_conclusion <- correlated_variables[
  nchar(correlated_variables$significantly_correlated_to)>0,
]
for (i in seq_len(nrow(correlated_variables_conclusion))) {
  print(
    paste0(
      "- ", correlated_variables_conclusion[i, "variable"], 
      " correlates with : ", correlated_variables_conclusion[i, "significantly_correlated_to"],
      "\n"
    )
  )
}

# you can make this example better by listing all pairs significantly correlated, 
# order pairs by alphabetic names, and remove duplicated info 
# (like Sepal.Length & Petal.Length and Petal.Length & Sepal.Length)
# etc 
```

